% !TEX root = ../main.tex
\chapter{Theoretical framework and empirical strategy}

This chapter outlines the theoretical framework used for our analysis. We begin by adapting a heterogeneous-firm trade model to the context of individual remittance behavior, enstablishing the theoretical foundations of how a migrant's transfer decisions can be influenced by economic shocks.

\section{A Remittance Model with Network Spillovers and Heterogeneous Migrants}

To analyze the impact of shocks on remittance flows, we adapt the logic of heterogeneous-firm trade models to the context of remittances. Specifically, we consider a model where migrants differ in income and altruism intensity, and where remitting requires paying both a fixed access cost and a variable cost proportional to the amount remitted. We then augment the model to allow migrants connected to the same origin municipality to respond to each other's transfers. Formally, consider migrant $i$, who resides in U.S. state $s$, originates from Mexican municipality $m$, and chooses remittances in period $t$.

We decompose origin-household resources into local non-remittance income and remittances received from migrants in other U.S. states. Perceived origin-household resources excluding migrant $i$'s own remittance are
\begin{equation}
    y^O_{imst}
    =
    \tilde{y}^O_{mt}
    +
    \sum_{s' \neq s} R_{s'mt},
\end{equation}
where $\tilde{y}^O_{mt}$ denotes local non-remittance income in municipality $m$, and $\sum_{s' \neq s} R_{s'mt}$ captures remittance inflows from migrants residing in U.S. states other than $s$. The migrant faces a utility maximization problem under a CES aggregator over local consumption and total resources available to the origin household:
\begin{equation}
    U_{imst}
    =
    \left[
    \alpha
    \left(c^L_{imst}\right)^{\frac{\sigma-1}{\sigma}}
    +
    (1-\alpha)
    \left(
    \phi_{im}
    \left[
    R_{imst}
    +
    \tilde{y}^O_{mt}
    +
    \sum_{s' \neq s} R_{s'mt}
    \right]
    \right)^{\frac{\sigma-1}{\sigma}}
    \right]^{\frac{\sigma}{\sigma-1}}.
\end{equation}
Here $\phi_{im} > 0$ denotes the altruism or obligation intensity parameter, analogous to the productivity parameter in \textcite{melitz2003}. This strictly positive parameter ensures that sending remittances can increase the migrant's utility by raising origin-household resources. The variable $c^L_{imst} \geq 0$ represents the migrant's consumption in the destination country, and $R_{imst} \geq 0$ is the amount sent in remittances by migrant $i$. The weight of local consumption is given by $\alpha \in (0,1)$, and the elasticity of substitution between local consumption and origin-household resources is given by $\sigma$. Origin-household total resources are therefore defined as $R_{imst} + y^O_{imst}$, which we treat as a single composite good. The presence of $y^O_{imst}$ introduces a role for both origin-country economic conditions and migration-network transfers: lower perceived origin resources increase the marginal utility of remittances, generating an insurance motive consistent with \textcite{yang_choi_2007}. 

\par 

The migrant's budget constraint is given by:
\begin{equation}
    c^L_{imst}
    +
    f_{mt}\mathbbm{1}\{R_{imst} > 0\}
    +
    (1 + \tau_{mt})R_{imst}
    \leq
    w_{imst},
\end{equation}
where $w_{imst} \geq 0$ is the migrant's wage, $f_{mt} \geq 0$ is a fixed remitting cost capturing the per-transaction expense of accessing a remittance channel, analogous to the fixed export cost in \textcite{melitz2003}, and $\tau_{mt} > -1$ is the variable cost which scales with the amount remitted, analogous to the iceberg cost structure in \textcite{melitz2003}. Notice that, even though $\tau_{mt}$ is typically positive to reflect transaction fees or exchange rate spreads, we allow for the possibility of negative variable costs, which could arise from the presence of remittance subsidies or favorable exchange rate regimes.

\par

Assuming that the migrant maximizes utility subject to the budget constraint and chooses to remit a positive amount ($R_{imst} > 0$), solving the Lagrangian yields the optimal remittance function:
\begin{equation}
    R^{*}_{imst}
    =
    \frac{
    A_{imst}
    \left[
    w_{imst}
    -
    f_{mt}
    +
    (1+\tau_{mt})
    \left(
    \tilde{y}^O_{mt}
    +
    \sum_{s' \neq s} R_{s'mt}
    \right)
    \right]
    }{
    1+(1+\tau_{mt})A_{imst}
    }
    -
    \left(
    \tilde{y}^O_{mt}
    +
    \sum_{s' \neq s} R_{s'mt}
    \right),
\end{equation}
where we define the parameter $A_{imst}$ as:
\begin{equation}
    A_{imst}
    \equiv
    \left(
    \frac{(1-\alpha)\phi_{im}}
    {\alpha(1+\tau_{mt})}
    \right)^{\sigma}.
\end{equation}
The term $A_{imst}$ represents the optimal ratio of origin-household resources to own local consumption, adjusted for the variable cost of remitting. It comprises the migrant's altruism intensity ($\phi_{im}$), the utility weight of own consumption ($\alpha$), the elasticity of substitution ($\sigma$), and the variable cost of remitting ($\tau_{mt}$).

\par

To understand the effect of destination-country economic shocks on remittance flows, we analyze changes in the migrant's wage ($w_{imst}$). The partial derivative of optimal remittances with respect to wage is strictly positive:
\begin{equation}
    \frac{\partial R^*_{imst}}{\partial w_{imst}}
    =
    \frac{A_{imst}}{1 + (1 + \tau_{mt})A_{imst}}
    >
    0.
\end{equation}
Therefore, through an income effect, the model predicts that a positive shock to the migrant's wage will increase remittances, while a negative shock will decrease them. The fixed cost $f_{mt}$ establishes an entry threshold: if the migrant's wage is not sufficiently high relative to the fixed cost, they will not send remittances at all. Additionally, the variable cost $\tau_{mt}$ reduces the marginal benefit of each dollar sent, dampening the sensitivity of remittances to wage fluctuations.

\par

The augmented model also delivers a spillover prediction. Holding the migrant's own wage and remitting costs fixed, remittances from other migrants enter the decision through perceived origin-household resources:
\begin{equation}
    \frac{\partial R^*_{imst}}{\partial R_{s'mt}}
    =
    -
    \frac{1}{1 + (1 + \tau_{mt})A_{imst}}
    <
    0
    \qquad
    \text{for } s' \neq s.
\end{equation}
A fall in remittances from one destination lowers perceived resources at origin and raises the optimal remittance from migrants in other destinations. For municipalities highly exposed to the shocked destination, this lowers origin-household resources and increases the marginal utility of additional transfers from migrants residing elsewhere. The model therefore provides a theoretical basis for compensatory migration-network spillovers: the same shock can generate a direct fall in remittances from the affected destination and an offsetting response from other connected destinations.

\par

The model delivers three empirical implications that guide the analysis. First, because optimal remittances are increasing in migrant income, a negative wage or income shock in the country where migrants reside should reduce remittances sent by affected migrants. Second, origin municipalities with greater pre-existing exposure to the affected migrant destination should experience larger declines in total remittance receipts, all else equal. This exposure effect is a reduced-form object: it captures the direct loss in transfers from the affected destination net of any adjustment by migrants in other destinations. Third, because remittances from other destinations enter the migrant's decision through perceived origin-household resources, a decline in transfers from the affected destination raises the marginal value of remittances from migrants in alternative destinations connected to the same municipality. The model therefore predicts compensatory responses from other high-sending destinations, especially for municipalities jointly exposed to the affected destination and those alternatives. These implications map directly into the empirical strategy below: we first estimate whether remittances fall from the shocked destination, then test whether municipal receipts fall more in places more exposed to that destination, and finally examine whether alternative corridors increase transfers relative to the shocked corridor among jointly exposed municipalities.
